Analysis of NLA17 Dissolved Gas Indicator

### 2023-03-22 manuscript/manuscript_file.Rmd
setting up manuscript infrastructure

---
title: "Widespread nitrous oxide undersaturation in U.S. lakes and reservoirs"
output: word_document
date: "2023-03-08"
---

@@

In this study we present the first national-scale survey **kasdflkjldfj** of N~2~O concentrations and emissions from lakes, reservoirs, and ponds, hereafter referred to as waterbodies, in the conterminous U.S. (CONUS). A generalized random-tessellation survey design (Stevens and Olsen, 2004) was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?]** sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week (Figure 1). Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate (See NLA website for all data?). `r round(((dg %>% dplyr::filter(n2o.src.snk == "sink") %>% {nrow(.)} / dg %>% distinct(site.id) %>% {nrow(.)}) * 100), 1)` percent of the sampled waterbodies were undersaturated in N2O and were therefore functioning as N2O sinks. We modeled the data using a bayesion hierchical approach and predicted the dissolved N2O concentration for all 224,916 CONUS waterbodies in the target population. At the population level, `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% pull()` percent (95% CI: `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% pull()` - `r all_predictions %>% group_by(.draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% pull()`) of CONUS waterbodies were functioning as N2O sink. The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N2O sinks (mean = `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% filter(WSA9 == "WMT") %>% select(estimate) %>% pull()`%, 95% CI: `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% filter(WSA9 == "WMT") %>% pull()` - `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% filter(WSA9 == "WMT") %>% pull()`) and the Northern Plains had the greatest (mean = `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(estimate = round(median(prop_sat), 3) * 100) %>% filter(WSA9 == "NPL") %>% select(estimate) %>% pull()`%, 95% CI: `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(LCL = round(quantile(prop_sat, probs = 0.025), 3) * 100) %>% filter(WSA9 == "NPL") %>% pull()` - `r all_predictions %>% group_by(WSA9, .draw) %>% summarise(prop_sat = sum(n2osat < 1) / length(.row)) %>% summarise(UCL = round(quantile(prop_sat, probs = 0.975), 3) * 100) %>% filter(WSA9 == "NPL") %>% pull()`)).

@@

# [3/9/2023] knitr can only find the file if placed in the root of the RStudio project.  Why can't it pull the figure from the manuscript_figures folder?
knitr::include_graphics("figure1.png")
```

>>>>

---
title: "Widespread nitrous oxide undersaturation in U.S. lakes and reservoirs"
output: word_document
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

@@

In this study we present the first national-scale survey of N~2~O concentrations and emissions from lakes, reservoirs, and ponds, hereafter referred to as waterbodies, in the conterminous U.S. (CONUS). A generalized random-tessellation survey design (Stevens and Olsen, 2004) was used to select a spatially-balanced and representative sample of `r dg %>% distinct(site.id) %>% summarize(n = n()) %>% pull()` **[weird, I have 1090, but NLA17 report says 1,005?]** sites from CONUS waterbodies that are least 1 hectare in area, at least 1 meter deep, have at least 0.1 hectare of open water, and have a minimum residence time of one week (Figure 1). Sample sites were distributed across all nine major CONUS ecoregions and varied widely in water chemistry, morphometry, watershed land-use, and climate (See NLA website for all data?). `r undersatN2oPercent` percent of the sampled waterbodies were undersaturated in N2O and were therefore functioning as N~2~O sinks (Figure 1). We modeled the data using a bayesion hierchical approach and predicted the dissolved N~2~O concentration for all 224,916 CONUS waterbodies in the target population. At the population level, `r propN2oSinkMean` percent (95% CI: `r propN2oSinkLower` - `r propN2oSinkUpper`) of CONUS waterbodies were functioning as N~2~O sinks. The Western Mountains ecoregion had the smallest proportion of waterbodies functioning as N~2~O sinks (mean = `r propN2oSinkMeanWM`%, 95% CI: `r propN2oSinkLowerWM` - `r propN2oSinkUpperWM`) and the Northern Plains had the greatest (mean = `r propN2oSinkMeanNPL`%, 95% CI: `r propN2oSinkLowerNPL` - `r propN2oSinkUpperNPL`).

@@

# [3/9/2023] knitr can only find the file if placed in the root of the RStudio project.  Why can't it pull the figure from the manuscript_figures folder?
#knitr::include_graphics("manuscript_figures/figure1.png")
```

### 2023-03-22 scripts/dgIndicatorAnalysis.Rmd
setting up manuscript infrastructure

## micromap: N2O saturation ratio continuous, extent, and emission rate
```{r}
# summarize means by WSA9 (posterior median for mean and 95% credible intervals)
# this is analogous to spsurvey "population estimates"
ecoreg_stats <- all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise( mean_n2o = mean(n2o),
             mean_n2oeq = mean(n2oeq),
             mean_sat = mean(n2osat),
             prop_sat = sum(n2osat < 1) / length(unique(.row)),
             mean_en2o = mean(e.n2o.mg.m2.d),
             fn2o_Mg = sum(f.n2o.Mg.y)) %>%
  summarise( post_med_n2o = round(mean(mean_n2o), 1),
             LCI_n2o = round(quantile(mean_n2o, probs = 0.025), 1),
             UCI_n2o = round(quantile(mean_n2o, probs = 0.975), 1),
             post_med_n2oeq = round(mean(mean_n2oeq), 2),
             LCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.025), 2),
             UCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.975), 2),
             post_med_sat = round(mean(mean_sat), 2),
             LCI_sat = round(quantile(mean_sat, probs = 0.025), 2),
             UCI_sat = round(quantile(mean_sat, probs = 0.975), 2),
             estimate = round(mean(prop_sat), 3),
             LCL = round(quantile(prop_sat, probs = 0.025), 3),
             UCL = round(quantile(prop_sat, probs = 0.975), 3),
             post_med_en2o = round(mean(mean_en2o), 3),
             LCL_en2o = round(quantile(mean_en2o, probs = 0.025), 3),
             UCL_en2o = round(quantile(mean_en2o, probs = 0.975), 3),
             post_med_fn2o = round(mean(fn2o_Mg), 3), 
             LCL_fn2o = round(quantile(fn2o_Mg, probs = 0.025), 3),
             UCL_fn2o = round(quantile(fn2o_Mg, probs = 0.975), 3),
             n = dplyr::n())

@@

ecoreg_stats$WSA9_NAME <- as.character(ecoreg_stats$WSA9)
ecoreg_stats <- ecoreg_stats %>% select(-WSA9)
str(ecoreg_stats)

@@

# N2O SATURATION AND UNDERSATURATION EXTENT
# Full code:  Use this to produce final linked micromaps of the variables you want
# export as pdf US letter landscape from Plots pane
# export as PNG Width 900 Height 968, maintain aspect ratio

@@

       print.file = "output/figures/mmapSatExtent.tiff", print.res = 300)

# N2O EMISSION AND FLUX
# Full code:  Use this to produce final linked micromaps of the variables you want
# export as pdf US letter landscape from Plots pane
# export as PNG Width 900 Height 968, maintain aspect ratio

>>>>

## MICROMAP: N2O saturation ratio continuous, extent, and emission rate

### Set up data
```{r}
# summarize means by WSA9 (posterior median for mean and 95% credible intervals)
# this is analogous to spsurvey "population estimates"
ecoreg_stats <- all_predictions %>%
  group_by(WSA9, .draw) %>%
  summarise( mean_n2o = mean(n2o),
             mean_n2oeq = mean(n2oeq),
             mean_sat = mean(n2osat),
             prop_sat = sum(n2osat < 1) / length(unique(.row)),
             mean_en2o = mean(e.n2o.mg.m2.d),
             fn2o_Mg = sum(f.n2o.Mg.y)) %>%
  summarise( post_med_n2o = round(mean(mean_n2o), 1),
             LCI_n2o = round(quantile(mean_n2o, probs = 0.025), 1),
             UCI_n2o = round(quantile(mean_n2o, probs = 0.975), 1),
             post_med_n2oeq = round(mean(mean_n2oeq), 2),
             LCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.025), 2),
             UCI_n2oeq = round(quantile(mean_n2oeq, probs = 0.975), 2),
             post_med_sat = round(mean(mean_sat), 2),
             LCI_sat = round(quantile(mean_sat, probs = 0.025), 2),
             UCI_sat = round(quantile(mean_sat, probs = 0.975), 2),
             estimate = round(mean(prop_sat), 3),
             LCL = round(quantile(prop_sat, probs = 0.025), 3),
             UCL = round(quantile(prop_sat, probs = 0.975), 3),
             post_med_en2o = round(mean(mean_en2o), 3),
             LCL_en2o = round(quantile(mean_en2o, probs = 0.025), 3),
             UCL_en2o = round(quantile(mean_en2o, probs = 0.975), 3),
             post_med_fn2o = round(mean(fn2o_Mg), 3), 
             LCL_fn2o = round(quantile(fn2o_Mg, probs = 0.025), 3),
             UCL_fn2o = round(quantile(fn2o_Mg, probs = 0.975), 3),
             n = dplyr::n())

@@

@@

### N2O SATURATION AND UNDERSATURATION EXTENT
```{r}
# Full code:  Use this to produce final linked micromaps of the variables you want
# export as pdf US letter landscape from Plots pane
# export as PNG Width 900 Height 968, maintain aspect ratio

@@

       print.file = "output/figures/mmapSatExtent.tiff", print.res = 300)
```

### N2O EMISSION AND FLUX
```{r}
# Full code:  Use this to produce final linked micromaps of the variables you want
# export as pdf US letter landscape from Plots pane
# export as PNG Width 900 Height 968, maintain aspect ratio
